---
title: "Goa style classifier"
author: "Tamas Nagy"
date: "10/24/2020"
output: 
  html_document:
   theme: paper
   code_download: true
   toc: true
   toc_float: true
editor_options: 
  chunk_output_type: console
---

Goa has several subgenres and some of them are difficult to tell apart. I want to build a classifier to predict genres based on music features.

```{r setup,	echo = TRUE, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(rvest)
library(spotifyr)
library(tidytext)
library(tidymodels)
library(igraph)
library(ggraph)

# Default ggplot theme
theme_set(theme_light())

# Setup parallel procesing
doParallel::registerDoParallel()

# Authenticate the spotify API through spotifyr.
spotify_id <- read_csv(here::here("spotify_client_id.txt"))

Sys.setenv(SPOTIFY_CLIENT_ID = spotify_id$client_id)
Sys.setenv(SPOTIFY_CLIENT_SECRET = spotify_id$secret)

access_token <- get_spotify_access_token()

```

## Create the dataset

The first task is to scrape the data that we can use for machine learning. Get data from <http://psytranceguide.com/> . There are genres here with examples, and what is more, links lead to spotify playlists. The spotifyr package makes it possible to retrieve track features for each playlist that we can use for classification.

Create a dataset that contains genre data and track features.

```{r, cache = TRUE}
# Scrape the homepage.
psytrance_link <- "http://psytranceguide.com/"
psytrance_page <- read_html(psytrance_link)

# Get the urls and other data from the psytrance guide
# I used the selectorGadget to find the proper css tags
goa_data <- 
  psytrance_page %>% 
  html_nodes("h2 a , .jouele-info-control-text")

# Create a table for all data (genres + playlist links)
goa_genres <-
  tibble(genre = map_chr(goa_data, html_text),
         playlist_url = html_attr(goa_data, "href")) %>% 
  extract(playlist_url, 
          into = "playlist_id", 
          regex = ".*playlist/(.*)\\?si.*") 

# Get all music features for all tracks for each playlist
features <- map_dfr(goa_genres$playlist_id, 
                    ~get_playlist_audio_features(playlist_uris = .x))


# Put playlist information and features together
goa <- 
  left_join(goa_genres, features, by = "playlist_id") %>% 
  hoist(track.album.artists, "name", .remove = TRUE) %>% 
  select(genre:playlist_name, 
         track_id = track.id, track_name = track.name, 
         artist_names = name,
         danceability:tempo) %>% 
  rowwise() %>% 
  mutate(artist_names = str_c(artist_names, collapse = ", ")) %>% 
  ungroup()

# Write data 
write_csv(goa, here::here("data/goa_tracks.csv"))

```

## Explore goa data

First let's see if there is enough variability in the features. I'm using a trick to calculate confidence intervals for the means by using a one-sample t.test with 0 as reference (could be any number as we are not interested in the p value).

```{r}
# Use the saved dataset so we don't have to reassemle the data.
goa <- 
  read_csv(here::here("data/goa_tracks.csv")) %>% 
  select(genre, track_name, danceability:tempo) %>% 
  mutate(genre = factor(genre))

# Put data into long format, calculate confidence intervals
goa_long <- 
  goa %>% 
  pivot_longer(danceability:tempo,
               names_to = "feature") %>% 
  group_nest(genre, feature) %>% 
  mutate(t_test = map(data, 
                      ~t.test(.x$value) %>% 
                      tidy())) %>% 
  unnest(c(t_test, data))
  
# Distributions
goa_long %>% 
  ggplot() +
  aes(x = value, fill = feature) +
  geom_histogram(show.legend = FALSE) +
  facet_wrap(~feature, scales = "free")

```

Let's order genres by features. Some features seem legit on face validity, although I have some difficulty in telling apart psyctrance genres. Also, I'm not fully aware of the meaning of all features.

```{r fig.height=15, fig.width=10}
goa_long %>% 
  mutate(genre = reorder_within(genre, estimate, feature)) %>%
  ggplot() +
  aes(x = estimate, y = genre, 
      xmin = conf.low, xmax = conf.high,
      color = feature) %>% 
  geom_pointrange(size = 0.5, show.legend = FALSE) +
  scale_y_reordered() +
  facet_wrap(~feature, scales = "free") +
  labs(title = "Average feature values by genre (95%CI)",
       subtitle = "Several features have high variability across genres that make features plausible to classify genres.",
       x = NULL, y = NULL)

```

## Classify goa subgenres using random forest

Create a random forest model for finding the defining features for the different subgenres. As we have a very little dataset, we are not splitting it into training and test set, but we use bootstraps to make the model more robust.

Bootstrap data, specify model and recipe.

```{r}
# Create bootstraps
set.seed(123)
goa_boot <- bootstraps(goa)

# Specify random forest model
rf_spec <- 
  rand_forest(trees = 1000) %>%
  set_mode("classification") %>%
  set_engine("ranger")

# Create workflow
goa_wf <- 
  workflow() %>%
  add_formula(genre ~ . -track_name) %>% 
  add_model(rf_spec)

```

Train model and evaluate. I seems like sensitivity is not great, that is a characteristic of multiclass classifications (too many categories to choose from). In overall, the model seems to have a 90% ROC AUC, that is better than expected.

```{r, cache = TRUE}
goa_rs <- 
  goa_wf %>% 
  fit_resamples(resamples = goa_boot,
                metrics = metric_set(roc_auc, specificity, sensitivity),
                control = control_resamples(save_pred = TRUE))

goa_rs %>% 
  collect_metrics()
```

Let's check the ROC curve for each class.

TODO

```{r eval=FALSE, include=FALSE}
goa_rs %>%
  collect_predictions() %>%
  group_by(id) %>%
  roc_curve(genre, .pred_Chillout:.pred_Tribal) %>%
  ggplot(aes(1 - specificity, sensitivity, color = id)) +
  geom_abline(lty = 2, color = "gray80", size = 1.5) +
  geom_path(show.legend = FALSE, alpha = 0.6, size = 1.2) +
  facet_wrap(~.pred_class, ncol = 5) +
  coord_equal()
```

## Explore which genres are easier to confuse with each other

Check the confusion matrix.

```{r}

goa_confusion <-
  goa_rs %>%
  collect_predictions() %>% 
  conf_mat(genre, .pred_class) 

goa_conf_perc <-
  goa_confusion$table %>% 
  as_tibble() %>% 
  group_by(Truth) %>% 
  mutate(all = sum(n),
         perc = n/all) 

goa_conf_perc %>% 
  ggplot() +
  aes(x = Truth, y = Prediction, fill = perc, 
      label = scales::percent(perc, accuracy = 1)) +
  geom_tile(show.legend = FALSE) +
  geom_text(size = 3) +
  labs(title = "Confusion matrix for the random forest model",
       subtitle = "Percent of correct classifications in the genre",
       x = "Truth")

# This shows the frequencies rather than percents (maybe remove?)
# goa_confusion %>% 
#   autoplot(type = "heatmap") +
#   ggtitle("Confusion matrix for the random forest model")
  
```

Let's visualize how easy it is to confuse certain genres with others by building a network graph!

```{r message=FALSE, warning=FALSE}

goa_conf_perc %>% 
  filter(Prediction != Truth & n > 0) %>% 
  select(Prediction, Truth, perc) %>% 
  graph_from_data_frame(directed = FALSE) %>% 
  ggraph() +
  geom_edge_link(aes(edge_width = perc, edge_alpha = perc)) +
  scale_edge_alpha(range = c(.1, .9)) +
  scale_edge_width(range = c(.01, 3)) +
  geom_node_text(aes(label = name), size = 4) +
  guides(fill = FALSE, alpha = FALSE, edge_alpha = FALSE, edge_width = FALSE) +
  theme_graph() +
  labs(title = "Confusion map of psytrance subgenres",
       subtitle = "The wider the connection between genre names, the easier to confuse")

    

```
